---
title: "Write-Up_Taxi-Analysis"
author: "Steven Chao, Tanaya Kavathekar, Madhuri Yadav, Amna Gul"
date: "`r Sys.Date()`"
output:
  html_document:
    code_folding: hide
    fig_height: 4.5
    fig_width: 7
    highlight: tango
    number_sections: yes
    theme: cosmo
    toc: yes
  pdf_document:
    toc: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(error = F)
#knitr::opts_chunk$set(echo = TRUE)
knitr::opts_chunk$set(include = F)
knitr::opts_chunk$set(warning = F)
```

```{r basicfcn, include=F}
# can add quietly=T option to the require() function
loadPkg = function(x) { if (!require(x,character.only=T, quietly =T)) { install.packages(x,dep=T,repos="http://cran.us.r-project.org"); if(!require(x,character.only=T)) stop("Package not found") } }


# This function calculates number of columns with type integer or double
count_numeric_columns = function(df) { numericc <- 0
doublec <- 0
for (col in colnames(df)){
  # check type integer
  if (typeof(df[,col]) == "integer"){
    numericc = numericc + 1
  }
  # check type double
  else {
    doublec = doublec + 1
  }
}
# create list of the return variables
  type_list <- list("num" = numericc, "dbl" = doublec) 
  return(type_list)
}
```

```{r load libraries, echo=FALSE}
# Load libraries
loadPkg("plyr")
loadPkg("dplyr")
loadPkg("ggplot2")
loadPkg("data.table")
loadPkg("nortest")
loadPkg("corrplot")
```

```{r read data files}
# Read in sample
unprocessed_data = read.csv("../Data/trips_zones_20000_v2.csv")

# Read in taxi zone name data
taxi_zones <- read.csv("../Data/taxi_zone_lookup.csv")
```

```{r Tip_Percentage}
# get percentage tip amount
unprocessed_data['per_tip'] = unprocessed_data['tip_amount'] / unprocessed_data['fare_amount']

# Print standard deviation and amount
print(sd(unprocessed_data$per_tip, na.rm = T))
print(mean(unprocessed_data$per_tip, na.rm = T))
```

```{r merge location descriptions}
# Merge taxi zone names to taxi cab data for pick up locations
unprocessed_data <- merge(x = unprocessed_data, y = taxi_zones, by.x = "PULocationID", by.y = "LocationID", all.x = TRUE)

# Delete unnecessary columns from merge
unprocessed_data <- subset(unprocessed_data, select = -c(Zone, service_zone))

# Change taxi zone name column
colnames(unprocessed_data)[colnames(unprocessed_data)=="Borough"] <- "Borough_pu"

# Merge taxi zone names to taxi cab data for drop off locations
unprocessed_data <- merge(x = unprocessed_data, y = taxi_zones, by.x = "DOLocationID", by.y = "LocationID", all.x = TRUE)

# Change taxi zone name column
colnames(unprocessed_data)[colnames(unprocessed_data)=="Borough"] <- "Borough_do"
```

```{r Pickup_time}
#Pickup column
#change factor type to datetime type
unprocessed_data$pickup_datetime <- as.POSIXct(as.character(unprocessed_data$tpep_pickup_datetime), tz = "", format="%m/%d/%Y %H:%M")
#extract only the time from datetime type
unprocessed_data$pickup_time <- format(unprocessed_data$pickup_datetime, format = "%H:%M")
#extract hours from the time extracted
unprocessed_data$pickup_hrs <- format(strptime(unprocessed_data$pickup_time,"%H:%M"),'%H')
#convert hrs column type to numeric
unprocessed_data$pickup_hrs <- as.numeric(unprocessed_data$pickup_hrs)
#convert numeric hrs type to pickup period
unprocessed_data$pickup_period <- ifelse(unprocessed_data$pickup_hrs >= 05 & unprocessed_data$pickup_hrs <= 11, "Morning",
            ifelse(unprocessed_data$pickup_hrs > 11 & unprocessed_data$pickup_hrs <= 16, "Afternoon",
            ifelse(unprocessed_data$pickup_hrs > 16 & unprocessed_data$pickup_hrs <= 21, "Evening", "Night")))
#convert pickup period from string type to factor
unprocessed_data$pickup_period <- as.factor(unprocessed_data$pickup_period)
#order the pickup_period column (this is of our interst in further analysis)
unprocessed_data$pickup_period <- ordered(unprocessed_data$pickup_period, levels = c("Morning", "Afternoon", "Evening", "Night"))
```

```{r Dropoff_time}
#Dropoff column
#change factor type to datetime type
unprocessed_data$dropoff_datetime <- as.POSIXct(as.character(unprocessed_data$tpep_dropoff_datetime), tz = "", format="%m/%d/%Y %H:%M")
#extract only the time from datetime type
unprocessed_data$dropoff_time <- format(unprocessed_data$dropoff_datetime, format = "%H:%M")
#extract hours from the time extracted
unprocessed_data$dropoff_hrs <- format(strptime(unprocessed_data$dropoff_time,"%H:%M"),'%H')
#convert hrs column type to numeric
unprocessed_data$dropoff_hrs <- as.numeric(unprocessed_data$dropoff_hrs)
#convert numeric hrs type to drop period
unprocessed_data$drop_period <- ifelse(unprocessed_data$dropoff_hrs >= 05 & unprocessed_data$dropoff_hrs <= 11, "Morning",
            ifelse(unprocessed_data$dropoff_hrs > 11 & unprocessed_data$dropoff_hrs <= 16, "Afternoon",
            ifelse(unprocessed_data$dropoff_hrs > 16 & unprocessed_data$dropoff_hrs <= 21, "Evening", "Night")))
#convert drop period from string type to factor
unprocessed_data$drop_period <- as.factor(unprocessed_data$drop_period)
#order the drop_period column(this is of our interst in further analysis)
unprocessed_data$drop_period <- ordered(unprocessed_data$drop_period, levels = c("Morning", "Afternoon", "Evening", "Night"))
```

```{r Trip_duration}
#Trip_duration
#calculation of trip duration in minutes = dropoff_time - pickup time
unprocessed_data$trip_duration <- difftime(unprocessed_data$dropoff_datetime, unprocessed_data$pickup_datetime, units = "mins")
#conversion of trip duration from minutes to numeric(this is of our interst in further analysis)
unprocessed_data$trip_duration <- floor(as.numeric(unprocessed_data$trip_duration))
str(unprocessed_data)
```

```{r tip-percent-summary2}
# Print structure
str(unprocessed_data)

# Print summary
summary(unprocessed_data)

# Print first five rows
head(unprocessed_data, n = 5)
```

```{r Outlier detection and missing values}
# fare amount in negative & credit card payments & trip_distance > 0 & VendorID == 4
processed_df <- unprocessed_data %>% filter((fare_amount > 0) & (payment_type == 1) &  (passenger_count < 7) & (trip_distance > 0)  & (VendorID != 4) & (trip_duration > 0) & (trip_duration <= 37))

# removing outliers:
# per_tip laying outside IQR range in box plot
while (length(boxplot(processed_df$per_tip)$out) != 0){
  dim(processed_df)
  per_tip_outliers <- boxplot(processed_df$per_tip, col = c("#0000FF"))
  processed_df <- processed_df[ !processed_df$per_tip %in% per_tip_outliers$out,]
  dim(processed_df)
}
```

Write a roughly 10-page (definitely no more than 4000 words) summary of the research and EDA process of your project. The summary should be prepared in R-markdown, and knitted into HTML. You may take some of the work in part B (such as graphs and results) to include here. They can overlap. This summary is to-be presented to your boss, your client, or to-be submitted for publication in journals. Potential area of topics to address in this summary may include:

#******EACH PERSON SHOULD WRITE ABOUT 600-700 WORDS!!!!!!*******
#*******MAKE SURE TO ADD SOME GRAPHS AND CHARTS************

# Introduction

Taxi cabs are an integral part of the New York City (NYC) experience. Widely recognized by their yellow color and checkered boxes, there are roughly 13,500 taxi cabs in service today, worth about $800,000 each (Devaraj & Patel, 2007). The popularity of taxi cabs have come under siege recently with the growing popularity and use of competing ride-hailing services such as Lyft and Uber. There are 80,000 of these cars in New York City, with 50,000 drivers serving 236 million passengers per year (DeBoard, 2019).

This study is relevant for a few reasons. Tips are an important part of the service industry, which includes taxicab drivers (Devaraj & Patel, 2017; Azar, 2007; Azar, 2010; Flynn, 2012). In general, tips average around 15% and generate billions of dollars each year (Azar, 2007). Thus, knowing the likelihood and the amount of tips have the potential to provide greater financial certainty for workers. In addition, despite New York law preventing taxi cab drivers from refusing service, service refusal still happens (Rivoli & Jorgensen, 2018). It is possible that drivers making assumptions about how likely a customer is to tip (Ayres, Vars, & Zakariya, 2005). Finally, taxi cab data can also provide valuable information on city life, city culture, human behavior, and various socioeconomic variables such as economic activity (Ferreira, Poco, Vo, Freire, & Silva, 2013).

# Literature Review

Tips can be defined as voluntary payments usually in the form of money—in addition to the obligatory transaction amount—from the customer to service worker who performs a service for them (Devaraj & Patel, 2017). There has been substantial research done in regard to tipping in the restaurant industry, mostly through empirical studies (Azar, 2007).

Tipping behavior can be explained by both economic and noneconomic factors (Devaraj & Patel, 2007). Economic reasons for tipping include providing an economic incentive for a higher quality of service (Flath, 2012) and wealth--although there are varying viewpoints amongst customers and waiters about the effect of a customer’s wealth on tip amount (Harris, 2012). Noneconomic factors include the idea of tipping being a social norm, where not tipping can cause a customer to feel guilty or embarrassed and where tipping can lead to positive feelings (Azar, 2010). Furthermore, tipping is also influenced by the fact that some customers recognize that their servers depend on tips to supplement their income.

Finally, gratitude or appreciation of service was cited as a reason (Azar, 2010). The extent to which the quality of service, however, affects the tip amount is unclear. Bodvarsson, Luksetich, and McDermott (2003), Bodvarsson and Gibson (1997), and Lynn and McCall (2000) through interviews all found that service quality greatly affects a customer’s decision to tip. Alternatively, Azar (2007) stresses the need to consider what is said during an interview versus what a customer actually does. While a customer may in principle agree with the fact that tips should depend on the quality of service, in an actual situation, the social norm of tipping may take precedence over that principle.

Another point of disagreement amongst researchers is the effect of weather on tip amounts. Flynn and Greenberg (2012) found that there is no statistically significant relationship between the two variables. They instead argue that tipping is an “institutional […] norm, rather than as a prosocial behavior that can be modulated by weather-induced changes in mood.”

## Taxicab Tipping

There exists a paucity of research on tipping in other industries, such as hotel and transportation (Azar, 2007; Devaraj & Patel, 2017. Devaraj and Patel (2017) argue that the factors previously discussed have less influence in a taxi setting given the standardized nature of the taxicab service and the minimal interaction between the driver and the passengers. Flath (2012) agrees, noting that the quality of the taxi service is not dependent on the driver.

## Taxicab Data Analysis

Various factors have been claimed to affect taxicab tipping. For instance, Devaraj and Patel (2017), who used NYC taxicab datasets, argue that taxicab tipping is associated with weather. While the relationship was small, it was statistically significant: The sunnier the weather, the greater amount of tipping. Moreover, Ayres et al. (2005) found that there is a racial disrimination in tipping. Conducting their study in New Hampshire, the researchers found that African-American drivers often received one-third of the tip received by white drivers. They also found that African-American and Hispanic passengers also tended to tip less than their white counterparts. They conclude by hypothesizing that this may be more of an indicator of wealth.

Other non-tipping work related to the NYC dataset have been conducted, especially in regards to location, For example, Zhan, Hasan, Ukkusuri, and Kamga (2013) use the dataset to estimate potential travel times between a pick up and drop off location. Neutens et al. (2012) use the dataset to conduct a spatio-temporal analysis of taxi trips in New York City. The authors design and introduce a new system that enables users to visualize taxi trips based on pick up and drop off locations.

Work has also been conducted in relation to supply and demand. Qian and Ukkusuri (2015) conducted a geographically-weighted regression on the dataset and found that commuting time is negatively related to taxi ridership, whereas subway accessibility is positively correlated to it. While they acknowledge that people with higher incomes theoretically are more able to afford taxis, they note the effect of median income varies by geographical location. Gonzales, Yang, Morgul, and Ozbay (2014) attempt to use socioeconomic data in conjunction with the taxicab dataset to model taxi demand, in particular from the NYC airports. Similar to Qian and Ukkusuri (2015), they conclude that accessibility, income, population, age, and number of jobs were strongly related to demand. Finally, Sun and McIntosh (2016) discover that taxicab ridership is lowest in the early mornings and the highest in the evenings. They also note that for short distances, trip time, distance, and fare are all linearly related; however, this relationship grows weaker at greater distances due to the influence of other factors.

## Relationship to Current Project

In this context, given the lack of research in taxicab tipping, this project seeks to address the gap in research by identifying if there are other factors other than weather that may affect the amount of tipping. If weather has the potential to affect behavior for tipping in taxicabs (Devaraj & Patel, 2017), there may be other factors, such as socio-economic variables, that may affect a person’s behavior and thus the amount they choose to tip.

As the literature review suggests, the provided dataset may provide some proxies for socio-economic variables. For instance, pick up and drop off locations may provide some context into a passenger's background for two main reasons. The first reason is quite simple: Passengers may use taxis to go from home to work and from work to home (Lee, Shin, & Park, 2008). Location can also impact the amount of tipping. This is because each location has its own set of customs and traditions and socioeconomic variables that affect tipping (Lynn, 2006). In this context, a person's wealth may also impact the amount that he or she tips to the taxi driver. While most of this analysis was done on a larger, regional scale, given the distinct cultures of each borough in New York, there may be microcosms of tipping culture found in the city that differ between each borough.Secondly, each individual has their own time-space geography. Some geographers have argued that people's time-spaces are often segregated--whether due to gender or race--which can affect their access to resources (Kwan, 1999; Neutens, Delafontaine, Scott, & De Maeyer, 2012; van Ham & Tammaru, 2015). Therefore, this may provide some context into how an individual lives and their background.

Finally, focusing on the variables provided in the dataset is beneficial because those variables are the ones the drivers “interact” with; they cannot see wealth or know a passenger’s background, but they can make inferences (such as on tip amount) based on appearances and their pick up and drop off locations (Ayres et al., 2005).


# Dataset Description
## How was the information gathered? - Madhuri
## What do we know about this dataset? - Madhuri
## preprocessing part - Madhuri- 

sampeling, provide some graphs and charts

The source dataset used for the project was taken from https://www1.nyc.gov/

Source: https://www1.nyc.gov/site/tlc/about/tlc-trip-record-data.page

  The New York City Taxi and Limousine Commission (TLC), was created in 1971.  The agency is responsible for licensing and regulating New York City's Medallion (Yellow) taxi cabs, for-hire vehicles (community-based liveries, black cars and luxury limousines), commuter vans, and paratransit vehicles. The datasets were collected and provided to the agency by technology providers authorized under the Taxicab & Livery Passenger Enhancement Programs (TPEP/LPEP). 
  
  The dataset is available for Yellow Taxi, Green Taxi and For-Hire Vehicle Records from 2009 till date. The total size of the complete dataset is around 10 GB. We have randomly selected 20000 observations from the latest dataset ie 2019 June from Yellow Taxi Records. due to hardware limitations, which have prevented us from analyzing the entire dataset.
  
  Below is the summary of the random sampled dataset (min, max, unique, column type) table:
  
  Data dictionary link
  
  We need to pre process data -Tanaya
  
  Normality
  Outlier
  Missing values
  Final EDA question with selected columns
  
  Below is the summary of the processed data:
  
  Before preprocessing we have 20,000 observations and 19 features. To bring the dataset to our EDA standards we performed preprocessing on it, thins involved column creations, outlier detection and treatment. 

  A normal distribution is often an assumption for many statistical analyses. Generally. raw tip amounts vary because the fare amounts vary. One factor that may not necessarily vary is tipping percentage. In the US, there is often a standardized percentage that a customer gives (for example, 15% at restaurants). We divided the fare amount by the tip amount to obtain a tip-to-fare ratio (hereinafter referred to as "tip ratio" or "tipping ratio").
  
  Analysis: The dataset provides a location ID that corresponds to a taxi zone in each of the five boroughs. These nominal variables do not provide much value in its integer format since we do not know the geographical locations of each location ID. We downloaded a taxi zone and ID dataset that provides the boroughs for each location ID. The dataset also indicates the specific neighborhoods within each borough. We merged that dataset to the taxi dataset to identify the borough for both pick up and drop off.
  
  Analysis: Pickup and Dropoff was in datetime format we changed it to categorical data with 4 values involving division of day into 4 major periods namely “Morning”, “Afternoon”, “Evening” or “Night”. Further we also calculated the trip duration in minutes based on these two columns, for our analysis.

To summarize, after removing outliers and creating new features we have 11881 observations and 33 features. 

# Discussion

## What are the limitations of the dataset? - Amna

Like every other data set, our data has also some limitations which may impact the results of our study.
<ul>
  <li>Hardware Limitations: The data that is publicly made available by the NYC Taxi and Limousine Commission dates back to only ten years. Initially we tried to incorporate all of this data for statistical testing but it was too large to be tested on our personal computers. Even data for a single year exceeded 12 Gbs causing our machines to crash or hang when trying to run it in R, so we narrowed it down to just 20,000 observations from the month of June 2019 for yellow taxi tip records (the latest available data). Since our observations span only a short period of time, therefore the results obtained from our research can not be generalized for longer period of times.
  <li>Non-Normality of Tipping data: Normally distributed data is a basic requirement when it comes to applying statistical tools, t-tests and the analysis of variance (ANOVA) etc. Our dependent variable i.e. Amount tipped is not exactly normally distributed. So we tried to address this issue in the following ways:
    <ul>
      <li>Removing extreme values and outliers that were causing our distribution to be abnormally skewed.
      <li>Central Limit Theorem (CLT) states that sample means of moderately large samples are often well-approximated by a normal distribution even if the data is not normally distributed. Our dataset contains a significant amount of observations thus qualifying it to be approximately normal under CLT. 
    </ul>
  <li>No Cash Tip data provided: The data available contains tip amounts payed through Credit Card mode only. There might be observations where Tip amount column contains a value of zero when in reality the driver of cab was presented with tip in cash mode but since it was not recorded in data collected by TLC so we cannot be sure and consequently rendering our data to be incomplete. 
  <li>Limited Time and resources: We chose to work on only four features affecting the variation in the tipping amount but there could be a plethora of other reasons too. For example weather. A bright sunny and breezy day would encourage short distance commuters to walk to their destinations instead of using cab service altogether or it might positively impact the mood of the taxi passengers inclining them to tip more. But limited time restricted us to work only on few of data points available.
  <li>Human Errors in Data set: Some of the columns in our data set are prone to human errors. For example, passenger count (the number of passengers in the vehicle) is driver entered value and  might have the potential of some observations being inaccurately recorded.
</ul>


## How did your question change, if at all, after Exploratory Data Analysis? - Tanaya

Initally our dependent variable was tip amount adn focused . We noticed that it didnot have normal distribution. Below is the initial distribution:

Initial question
-not normal, so we changed to tip amount
-because only credit card tips, we changed our question to focus only on credit card tips

-put updated research questions here

from COMBINED file

QUESTION: For those who pay tip by credit card, what factors influence the amount of tip given?
Should you pay tip to the taxi driver? If yes, how much?

Analyse the factors which takes place to determine the tip amount paid to the driver

Why?
Benefits?
Use?

Hypothesis:
Factors that can affect tip amount:
1. Vendor service
2. Driver service
3. Location :  (generous locations)
4. duration of day 
5. distance
6. count of passengers
7. weather


## What additional information would be beneficial? - Amna

The results obtained by our team can be significantly improved and made more reliable if we were able to gather more features to consider. For instance:
<ul>
  <li>Driver ratings: If TLC was able to gather data on how well-rated each Yellow Cab driver is in NYC based on feedback provided by passenger(s). This rating system itself could involve various factors like cleanliness level of vehicle interior, friendliness/politeness level of driver etc. thus allowing us to find the correlation between amount tipped to good rated drivers versus not so good rated drivers.
  <li>Weather data: Common sense tells us that rainy or snowy weather is going to encourage more people to opt for cab hailing services than on normal days. From that data we could have applied statistical techniques to infer whether the change in season also affects the tipping behavior of the passengers or not. 
  <li>Geographical Coordinates: The availability of geographical coordinates data would not only have led us to pinpoint the exact localities/neighborhoods to figure out the tipping intention based on different demographics but also projected this data in the form of visually appealing maps.
  <li>Gender Dynamics, Personality type, Mood: These factors due to their qualitative nature receive limited attention as predictor of tipping behavior but we believe that can also serve to play an important part in tipping behavior.
</ul>

# Conclusion
## Based on EDA can you begin to sketch out an answer to your question? - EVERYONE ON SUNDAY

PUT ONLY THE RELEVANT GRAPHS HERE, ONE FOR EACH OF YOUR VARIABLES

For tip percentage and location, after running the ANOVA test, we reject the null hypothesis that the means are the same and say the means are statistically different at a significance level of 0.05. While the means are overall not the same, the following pick up pairs have significant differences in tipping percentage: Manhattan and Brooklyn, and Queens and Manhattan. For drop off pairs, Manhattan and Bronx, Manhattan and Brooklyn, and Manhattan and Queens are significant. The graphs for tip percentage and location are shown below:

```{r tipper_plot, include = T, echo = T}
# Plot tip ratio by pickup location
ggplot(data = processed_df, aes(x = Borough_pu, y = per_tip, fill = Borough_pu)) + ggtitle("Tip Ratio by Pickup Location") + geom_boxplot() + xlab("Location") + ylab("Tip Ratio")

# Plot tip ratio by dropoff location
ggplot(data = processed_df, aes(x = Borough_do, y = per_tip, fill = Borough_do)) + ggtitle("Tip Ratio by Dropoff Location") + geom_boxplot() + xlab("Location") + ylab("Tip Ratio")
```

AMNA SUMMARY/GRAPH

MADHURI SUMMARY/GRAPH

TANAYA SUMMARY/GRAPH

next steps:
Linear regression; look at pearson correlation; then build linear model


# Sources

Ayres, I., Vars, F. E., & Zakariya, N. (2005). To Insure Prejudice: Racial Disparities in Taxicab Tipping. Yale Law Journal, 114, 1613–1674. doi: 10.2139/ssrn.401201

Azar, O. H. (2007). The Social Norm of Tipping: A Review. Journal of Applied Social Psychology, 37(2), 380–402. doi: 10.1111/j.0021-9029.2007.00165.x

Azar, O. H. (2010). Tipping Motivations and Behavior in the U.S. and Israel. Journal of Applied Social Psychology, 40(2), 421–457. doi: 10.1111/j.1559-1816.2009.00581.x

DeBord, M. (2019, April 1). Congestion pricing could mark the beginning of the end of New York's famous yellow taxis. Retrieved October 14, 2019, from https://www.businessinsider.com/congestion-pricing-end-of-new-yorks-famous-yellow-taxis-2019-4.

Devaraj, S., & Patel, P. C. (2017). Taxicab tipping and sunlight. PLoS One, 12(6). doi: 10.1371/journal.pone.0179193

Douriez, M., Doraiswamy, H., Freire, J., & Silva, C. T. (2016). Anonymizing NYC Taxi Data: Does It Matter? Proceedings - 3rd IEEE International Conference on Data Science and Advanced Analytics, 140–148. doi: 10.1109/DSAA.2016.21

Ferreira, N., Poco, J., Vo, H. T., Freire, J., & Silva, C. T. (2013). Visual Exploration of Big Spatio-Temporal Urban Data: A Study of New York City Taxi Trips. IEEE Transactions on Visualization and Computer Graphics, 19(12), 2149–2158. doi: 10.1109/TVCG.2013.226

Flath, D. (2012). Why Do We Tip Taxicab Drivers? Japanese Economy, 39(3), 69–76. doi: 10.2753/JES1097-203X390304

Flynn, S. M., & Greenberg, A. E. (2012). Does Weather Actually Affect Tipping? An Empirical Analysis of Time-Series Data. Does Weather Actually Affect Tipping? An Empirical Analysis of Time-Series Data, 42(3), 702–716. doi: Does Weather Actually Affect Tipping? An Empirical Analysis of Time-Series Data

Gonzales, E. J., Yang, C., Morgul, E. F., & Ozbay, K. (2014). Modeling Taxi Demand with Gps Data from Taxis and Transit. Modeling Taxi Demand with GPS Data from Taxis and Transit. Mineta National Transit Research Consortium. Retrieved from https://transweb.sjsu.edu/sites/default/files/1141-modeling-taxi-demand-gps-transit-data.pdf

Harris, M. B. (1995). Waiters, Customers, and Service: Some Tips About Tipping. Journal OfApplied Social Psychology, 25(8), 725–744. doi: 10.1111/j.1559-1816.1995.tb01771.x

Kwan, M. (1999). Gender and Individual Access to Urban Opportunities: A Study Using Space–Time Measures. The Professional Geographer, 51, 211–227. doi: 10.1111/0033-0124.00158

Lee, J., Shin, I., & Park, G. (2008). Analysis of the Passenger Pick-Up Pattern for Taxi Location Recommendation. 2008 Fourth International Conference on Networked Computing and Advanced Information Management. doi: 10.1109/NCM.2008.24

Lynn, W. M. (2004). Tipping in Restaurants and Around the Globe: An Interdisciplinary Review. Handbook of Contemporary Behavioral Economics: Foundations and Developments. doi: https://ssrn.com/abstract=465942

Neutens, T., Delafontaine, M., Scott, D. M., & De Maeyer, P. (2012). An analysis of day-to-day variations in individual space-time accessibility. Journal of Transport Geography, 23, 81–91. doi: 10.1016/j.jtrangeo.2012.04.001

Qian, X., & Ukkusuri, S. (2015). Spatial variation of the urban taxi ridership using GPS data. Applied Geography, 59, 31–42. doi: 10.1016/j.apgeog.2015.02.011

Rivoli, D., & Jorgensen, J. (2018, July 31). City vows to crack down on taxis refusing service as it also looks to cap Uber. Retrieved October 14, 2019, from https://www.nydailynews.com/news/politics/ny-pol-tax-uber-refuse-service-20180731-story.html.

Sun, H., & McIntosh, S. (2016). Big Data Mobile Services for New York City Taxi Riders and Drivers. 2016 IEEE International Conference on Mobile Services, 57–64. doi: 10.1109/MoSb.2S0e1r6v.128016.19

van Ham, M., & Tammaru, T. (2015). New perspectives on ethnic segregation over time and space. A domains approach. Urban Geography, 37(7), 953–962. doi: 10.1080/02723638.2016.1142152

Zhan, X., Hasan, S., Ukkusuri, S., & Kamga, C. (2013). Urban link travel time estimation using large-scale taxi data with partial information. Transportation Research Part C, 33, 37–49. doi: 10.1016/j.trc.2013.04.001