---
title: "WRITE-UP_Taxi-Analysis"
author: "Steven Chao"
date: "10/12/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Write a roughly 10-page (definitely no more than 4000 words) summary of the research and EDA process of your project. The summary should be prepared in R-markdown, and knitted into HTML. You may take some of the work in part B (such as graphs and results) to include here. They can overlap. This summary is to-be presented to your boss, your client, or to-be submitted for publication in journals. Potential area of topics to address in this summary may include:
##What do we know about this dataset? - Madhuri


##What are the limitations of the dataset? - Amna


##How was the information gathered? - Madhuri


##What analysis has already been completed related to the content in your dataset? -Steven


##How did the research you gathered contribute to your question development? - Steven


##What additional information would be beneficial? - Amna


##How did your question change, if at all, after Exploratory Data Analysis? - Tanaya

```{r cars}
summary(cars)
```


##Based on EDA can you begin to sketch out an answer to your question? - EVERYONE